Pattern matching for\\ $(213,231)$-avoiding permutations

### 2015-06-16 article.tex
modif proof of proposition 3 plus typos

%first
\DeclareMathOperator{\firsta}{L}
\newcommand{\first}[2]{\firsta_{{#1}}({#2})}

\DeclareMathOperator{\firstia}{LI}
\newcommand{\firsti}[2]{\firstia_{{#1}}({#2})}

\DeclareMathOperator{\factora}{F}
\newcommand{\factor}[2]{\factora_{{#1}}({#2})}

@@

	This paper is organized as follows.
	In Section~\ref{section:Definitions} the needed definition are presented.
	Section~\ref{section:both are (231,213)-avoiding} is devoted to presenting
	an online linear-time algorithm in case both the pattern and the target
	permutation are $(213,231)$-avoiding,
	whereas Section~\ref{section:sigma only avoid 231 and 213} focuses on the case
	where the pattern only is $(213,231)$-avoiding.
	In Section~\ref{section:bivincular} we give a polynomial-time algorithm
	for $(213,231)$-avoiding bivincular patterns.
	In Section~\ref{section:LCS} we consider the problem of finding the longest
	$(213,231)$-avoiding pattern in a permutation.
	In Section~\ref{section:LCS common 2 permutations} we give a
	polynomial-time algorithm for computing the longest
	$(213,231)$-avoiding pattern common to two permutations.

@@

\section{Both $\pi$ and $\sigma$ are $(213,231)$-avoiding}
\label{section:both are (231,213)-avoiding}

@@

\begin{proposition}
	Let $\pi$ and $\sigma$ be two $(213,231)$-avoiding permutations.
	One can decide whether $\pi$ contains $\sigma$ in linear-time.
\end{proposition}

\begin{proof}
According to Lemma~\ref{lemma:MatchStripeToPermutation} the problem reduces
to deciding whether $s_\sigma$ occurs as a subsequence in $s_\pi$.
A straightforward greedy approach solves this issue in linear-time.
\qed
% To solve the permutation pattern for a permutation in $\AV_n(213,231)$,
% we find a find a match of the stripe of the pattern in the stripe of the text.
% This can be done linearly by a greedy algorithm,
% by matching a step whenever it is possible.
% And thank to the definition of the definition of a stripe in
% Corollary~\ref{corollary:minmaxelement},
% we do not have to compute the stripe in advance.
% Therefore we have an linear (online) algorithm.
\end{proof}

Observe that, according to Corollary~\ref{corollary:minmaxelement},
the effective construction of the two stripes is not required for the above proposition,
and hence we have an online algorithm.

@@

\section{$\sigma$ only is $(231,213)$-avoiding}
\label{section:sigma only avoid 231 and 213}

This section focuses on the pattern containment problem
in case only the pattern $\sigma$ avoids $231$ and $213$.
We need to consider a specific decomposition $\factora$ of $\sigma$ into factors.
For $\sigma \in S_k$, the \emph{factorization}
$\sigma = \factor{\pmotif}{p} \factor{\pmotif}{p-1} \ldots \factor{\pmotif}{1}$
is defined as follows:
(i) $\sigma[k]$ is the last letter of $\factor{\pmotif}{1}$,
(ii) for every $1 \leq i < k$,
if $s_\sigma[i] = s_\sigma[i+1]$ then
$\sigma[i]\sigma[i+1]$ is part of the same factor.
For every factor $\factor{\pmotif}{j}$ of the decomposition, we let
$\first{\pmotif}{j}$ stand for the leftmost letter of $\factor{\pmotif}{j}$,
and $\firsti{\pmotif}{j}$ stand for the index in $\sigma$
of the leftmost letter of $\factor{\pmotif}{j}$.
For example,
for $\sigma = 981237654$,
we have
$s_\sigma = ddaaaddd$ and hence
$\sigma = \factor{\pmotif}{3} \factor{\pmotif}{2} \factor{\pmotif}{1}$ with
$\factor{\pmotif}{1} =7654$, $\factor{\pmotif}{2} = 123$ and $\factor{\pmotif}{3} = 98$.
Furthermore,
$\first{\pmotif}{1} = 7$, $\first{\pmotif}{2} = 1$, $\first{\pmotif}{3} = 9$,
$\firsti{\pmotif}{1} = 6$, $\firsti{\pmotif}{2} = 3$ and $\firsti{\pmotif}{3} = 1$.

\begin{remark}
A Factor is either an increasing or a decreasing subsequence.
\end{remark}

\begin{lemma}
\label{lemma:whereIsMax}
Given a permutation $\sigma \in \AV(213,231)$ and a suffix of its 
decomposition $\factor{\sigma}{i}\factor{\sigma}{i-1}\ldots\factor{\sigma}{1}$, 
if $\firsti{\sigma}{i}$ is an ascent (resp. descent) element then the maximal 
(resp. minimal) element of $\factor{\sigma}{i}\factor{\sigma}{i-1}\ldots\factor{\sigma}{1}$ is 
$\first{\sigma}{i-1}$.
\end{lemma}

@@

we now define the set
$\SET{\pmotif}{\ptext}{i}{j}$ as follows:
given a subsequence $s$ of $\ptext[j:]$,
$s \in \SET{\pmotif}{\ptext}{i}{j}$ if and only if
$s$ is an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$
and the element$\first{\pmotif}{i}$ occurs in $\ptext[j]$.

\begin{lemma}
Let $\pmotif$ be a permutation 
with $\factor{\pmotif}{i}$ as an ascent (respectively descent) factor,
$s$ a subsequence such that, $s \in \SET{\pmotif}{\ptext}{i}{j}$ and $s[\first{\pmotif}{i-1}]$ is minimal (resp. maximal). 
For all subsequence 
$s' \in \SET{\pmotif}{\ptext}{i}{j}$ and for all subsequence $t$ of $\ptext$, such that $t=t's'$, if $t$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ then the subsequence $t's$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$.
\end{lemma}

@@

\begin{corollary}
Let $\pmotif$ be a permutation 
with $\factor{\pmotif}{i}$ as an ascent (respectively descent) factor,
and let $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that $s[\first{\pmotif}{i-1}]$ is minimal (resp. maximal). 
Those following statements are equivalent :
\begin{itemize}
	\item There exists an 
	occurrence of $\pmotif$ in $\ptext$ with element $\first{\pmotif}{i}$ occurs in $\ptext[j]$.
	\item There exists a subsequence $t$ of $\ptext[1:j-1]$  such that $ts$ is an occurrence of $\pmotif$ in $\ptext$ with element $\first{\pmotif}{i}$ occurs in $\ptext[j]$. 
\end{itemize}
\end{corollary}

@@

	\text{The maximal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ which is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & 
		\text{ an ascent element } \\	
	\text{ and that minimize the maximal value} & \\
	\text{ and $\first{\sigma}{i}$ occurs in $\ptext[j]$} & \\

@@

	\text{The minimal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ which is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & \text{ a descent element } \\	
	\text{ and that maximize the minimal value} & \\
	\text{ and $\first{\sigma}{i}$ occurs in $\ptext[j]$} & \\
\end{cases}
$$

@@

\noindent\textbf{BASE :}
\begin{align*}
\LM{\pmotif}{\ptext}{1}{j} 
&=
\begin{cases}
		\min_{j<j'}\{\infty\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LIS(j,j',\ptext[j'])  $}\}&\text{ an ascent element } \\
		&\\
		\max_{j<j'}\{0\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LDS(j,j',\ptext[j']) $}\}&\text{ a descent element }\\
\end{cases}
\\
\intertext{\textbf{STEP :}}
\LM{\pmotif}{\ptext}{i}{j} 
&=
\begin{cases}
	\min \{\infty\} \cup  \AF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is an ascent element}\\
	\max \{0\} \cup  \DF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is  a descent element}\\
\end{cases}
\end{align*}
where $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ are the sets 
of elements such that if  
$\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$ or if $\ptext[k] \in \DF{\pmotif}{\ptext}{i}{j}$ 
then there exists an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$ and the element 
$\first{\pmotif}{i-1}$ occurs in element $\ptext[k]$. 
Given that one has the set $\SET{\pmotif}{\ptext}{i-1}{k}$, 
to decide if $\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$, 
one needs to find an occurrence $t$ of $\factor{\pmotif}{i}$ in $\ptext[j:k-1]$ 
such there exists $s \in \SET{\pmotif}{\ptext}{i-1}{k}$ and $ts$ is an occurrence of 
$\pmotif[\firsti{\pmotif}{i}:]$. 
Formally we define $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ as follows:
\begin{align*}
\AF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq 0$ and} 
\\
&\qquad 
\text{$|\factor{\pmotif}{i}| \leq LIS_{\ptext}(j,j'-1,\first{\pmotif}{i-1})\}$} 
\\
\DF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq \infty$ and}
\\
&\qquad
\text{$|\factor{\pmotif}{i}| \leq LDS_{\ptext}(j,j'-1,\first{\pmotif}{i-1})\}$}
\end{align*}
where $LIS_{\ptext}(i,j,k)$ (resp. $LDS_{\ptext}(i,j,k)$) is the longest increasing 
(resp. decreasing) sequence in $\ptext$ starting at $i$ and ending at $j$,
with every element of this sequence
smaller (resp. bigger) than $k$.
$LIS_{\ptext}$ and $LDS_{\ptext}$ can be computed in 
$O(|\ptext|^2\log(\log(|\ptext|)))$ (see \cite{Bespamyatnikh00enumeratinglongest}).

For the base case,
one is looking for an occurrence for the first factor.
If the factor is an ascent (respectively descent) factor, then
one has to find an increasing (resp. decreasing) subsequence
in the text of same size or longest that
the size of the first factor.
To find the "safest" solution
one must assure that the right most element
of the sequence is minimal (resp. maximal).\\
For the induction, it follows the same idea
except that one must assure that
every element of the subsequence is larger (resp. smaller)
than the maximal (resp. minimal) element of the rest of the occurrence which is given by the inductive problem of
the previous factor.

There exists a occurrence of $\pmotif$ in $\ptext$ if and only if
there exists a $LM(n,i)\neq 0$ and $LM(n,i)\neq \infty$ for $1 \leq i \leq |\ptext|$,
with $n$ the number of factor in $\pmotif$.
Moreover the basic case can be computed in $O(n^2)$
and the induction on $O(kn^2)$ time and space.
\qed
\end{proof}

@@

This section is devoted to presenting a polynomial-time algorithm for deciding whether
a $(213,231)$-avoiding bivincular pattern occurs in a permutation.
We start by presenting a property of bivincular motifs avoiding both $213$ and $231$.

\begin{lemma}
\label{lemma:ascentDescentAscent}
If $\overline{\sigma[i]\sigma[j]}$ (thus $\sigma[j]=\sigma[i]+1$) and 
if $\sigma[i]$ is an ascent (respectively decent) element and $\sigma[i]+1$ is an ascent 
(resp. decent) element then 
$i<j$ (resp. $j>i$) and for every $k$, $i<k<j$ (resp. $j>k>i$), $\pi[k]$ is a descent 
(resp. ascent) element.    
\end{lemma}

\begin{proof}
Suppose that there exists $k$, $i<k<j$,
such that $\pmotif[k]$ is ascent. Ascent elements are increasing  so $\sigma[i]<\sigma[k]<\sigma[j]$ which is in contradiction that with $\sigma[j]=\sigma[i]+1$.
We use a similar argument if $\sigma[i]$ is a descent element
\qed
\end{proof}

\begin{proposition}
	\label{Proposition:bivincular pattern}
		Let $\sigma$ be a bivincular motif avoiding $213$ and $231$ of length $k$
		and $\pi \in S_n$.
		One can decide in $O(kn^5)$ time
		and $O(kn^5)$ space if $\sigma$ occurs in $\pi$.
\end{proposition}

\begin{proof}
We consider the following problem :
Given a bivincular motif $\pbmotif$ with $\pmotif \in \Av{n_\pmotif}{231,213} $, and a text $\ptext \in \Av{n_\ptext}{231,213}$.
$$
\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}=
\begin{cases}
	true 	& \text{If $\pbmotif[i:]$ occurs in $\ptext[j:]$}\\
			& \text{with every element is in $[\lb,\ub]$}\\
			& \text{and element $\pmotif[i]$ occurs in $\ptext[j]$ }\\
	false 	& otherwise\\
\end{cases}
$$

$\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}$ is closed under induction.
It can be solved by means by the following relations:

@@

The base case finds an occurrence for the right most element of the motif. If the last element does not have any restriction on position and on value, then $\PM{\pbmotif}{\ptext}{\lb}{\ub}{n_\pmotif}{j}$ is true if and only if $\pmotif[n_\pmotif]$ can occur in $\ptext[j]$. This condition is checked by assuring that $\ptext[j] \in [\lb,\ub]$.

@@

We consider 3 cases for the problem $\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}$ :
\begin{itemize}
	\item If $\ptext[j] \notin [\lb,\ub]$ then :
	$$
	\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j} = false
	$$

	this is straightforward with the definition. If $\ptext[j] \notin [\lb,\ub]$ then it can not be part of an occurrence of  $\pmotif[i:]$ in $\ptext[i:]$ with every element is in $[\lb,\ub]$.

@@

	\end{cases}
	$$	
	Remark that $\pmotif[i]$ can occur in $\ptext[j]$ because $\ptext[j] \in [\lb,\ub]$.
	Thus if $\pmotif[i+1:]$ occurs in $\ptext[j+1:]$ with every element in $[\pmotif[i]+1,\ub]$ then 
 	$\pmotif[i:]$ occurs in $\ptext[j:]$  	
 	which correspond to know if there exists $k$, $j<k$ such that
	$\PM{\pbmotif}{\ptext}{\ptext[j]+1}{\ub}{i+1}{k}$ is true.

@@

	$$
	\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}=
	\begin{cases}
			\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}
				& \text{if $\pmotif[i]$ is not underlined } \\
				& \text{ and $\pmotif[i]$ is not overlined} \\
			\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}
				& \text{if $\pmotif[i]$ is not underlined } \\
				& \text{ and $\overline{\pmotif[i](\pmotif[i]+1)}$ or ${\pmotif[i]}^\urcorner$}\\
				& \text{ and $\ptext[j]=\ub$} \\
			\PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{j+1}
				& \text{if $\underline{\pmotif[i]\pmotif[i+1]}$ } \\
				& \text{ and $\overline{\pmotif[i](\pmotif[i]+1)}$ or ${\pmotif[i]}^\urcorner$}\\
				& \text{ and $\ptext[j]=\ub$} \\
			\PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{j+1}
				& \text{if $\underline{\pmotif[i]\pmotif[i+1]}$ } \\
				& \text{ and $\pmotif[i]$ is not overlined} \\
			false & \text{otherwise}
	\end{cases}
	$$	
	We can show symmetrically as the last case that 
	$\pmotif[i:]$ occurs in $\ptext[j:]$
	with $\sigma[i]$ occurs in $\pi[j]$.
	 starting at element $j$ if and only if $\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}$.

@@

\begin{proposition}
Given two permutations $\pi_1$ and $\pi_2$,
one can compute 
the longest common subsequence
avoiding (231,213) in $O(|\pi_1|^3|\pi_2|^3)$ time and space.
\end{proposition}

>>>>

\DeclareMathOperator{\firstia}{first}
\newcommand{\firsti}[2]{\firstia_{{#1}}({#2})}

\DeclareMathOperator{\factora}{factor}
\newcommand{\factor}[2]{\factora_{{#1}}({#2})}

%first
\DeclareMathOperator{\firsta}{first}
\newcommand{\first}[2]{\firsta(\factor{#1}{#2})}

@@

	This paper is organized as follows.
	In Section~\ref{section:Definitions} the needed definition are presented.
	Section~\ref{section:both are (213,231)-avoiding} is devoted to presenting
	an online linear-time algorithm in case both the pattern and the target
	permutation are $(213,231)$-avoiding,
	whereas Section~\ref{section:sigma only avoid 231 and 213} focuses on the case
	where the pattern only is $(213,231)$-avoiding.
	In Section~\ref{section:bivincular} we give a polynomial-time algorithm
	for $(213,231)$-avoiding bivincular patterns.
	In Section~\ref{section:LCS} we consider the problem of finding the longest
	$(213,231)$-avoiding pattern in a permutation.
	In Section~\ref{section:LCS common 2 permutations} we give a
	polynomial-time algorithm for computing the longest
	$(213,231)$-avoiding pattern common to two permutations.

@@

\section{Both $\pi$ and $\sigma$ are $(213,231)$-avoiding}
\label{section:both are (213,231)-avoiding}

@@

\begin{proposition}
	\label{Proposition:both permutations are avoiding}
	Let $\pi$ and $\sigma$ be two $(213,231)$-avoiding permutations.
	One can decide whether $\pi$ contains $\sigma$ in linear time.
\end{proposition}

\begin{proof}
According to Lemma~\ref{lemma:MatchStripeToPermutation} the problem reduces
to deciding whether $s_\sigma$ occurs as a subsequence in $s_\pi$.
A natural greedy approach solves this issue in linear-time.
\qed
% To solve the permutation pattern for a permutation in $\AV_n(213,231)$,
% we find a find a match of the stripe of the pattern in the stripe of the text.
% This can be done linearly by a greedy algorithm,
% by matching a step whenever it is possible.
% And thank to the definition of the definition of a stripe in
% Corollary~\ref{corollary:minmaxelement},
% we do not have to compute the stripe in advance.
% Therefore we have an linear (online) algorithm.
\end{proof}

Observe that, according to Corollary~\ref{corollary:minmaxelement},
the effective construction of the two stripes is not required in stretch 
of algorithm given for the proof in the proposition \ref{Proposition:both permutations are avoiding},
and hence we have an online algorithm.

@@

\section{$\sigma$ only is $(213,231)$-avoiding}
\label{section:sigma only avoid 231 and 213}

This section focuses on the pattern containment problem
in case only the pattern $\sigma$ avoids $231$ and $213$.
We need to consider a specific decomposition $\factora$ of $\sigma$ into factors.
For $\sigma \in S_k$, the \emph{factorization}
$\sigma = \factor{\pmotif}{p} \factor{\pmotif}{p-1} \ldots$ $\factor{\pmotif}{1}$
is defined as follows:
(i) $\sigma[k]$ is the last letter of $\factor{\pmotif}{1}$,
(ii) for every $1 \leq i < k$,
if $s_\sigma[i] = s_\sigma[i+1]$ then
$\sigma[i]\sigma[i+1]$ is part of the same factor.
For every factor $\factor{\pmotif}{j}$ of the decomposition, we let
$\first{\pmotif}{j}$ stand for the leftmost letter of $\factor{\pmotif}{j}$,
and $\firsti{\pmotif}{j}$ stand for the index in $\sigma$
of the leftmost letter of $\factor{\pmotif}{j}$.
For example,
for $\sigma = 981237654$,
we have
$s_\sigma = ddaaaddd$ and hence
$\sigma = \factor{\pmotif}{3} \factor{\pmotif}{2}$ $ \factor{\pmotif}{1}$ with
$\factor{\pmotif}{1} =7654$, $\factor{\pmotif}{2} = 123$ and $\factor{\pmotif}{3} = 98$.
Furthermore,
$\first{\pmotif}{1} = 7$, $\first{\pmotif}{2} = 1$, $\first{\pmotif}{3} = 9$,
$\firsti{\pmotif}{1} = 6$, $\firsti{\pmotif}{2} = 3$ and $\firsti{\pmotif}{3} = 1$.

\begin{remark}
All the elements of a factor are all increasing elements or all decreasing elements
\end{remark}

\begin{lemma}
\label{lemma:whereIsMax}
Given a permutation $\sigma \in \AV(213,231)$ and a suffix of its 
decomposition $\factor{\sigma}{i}\factor{\sigma}{i-1}\ldots$$\factor{\sigma}{1}$, 
if $\firsti{\sigma}{i}$ is an ascent (resp. descent) element then the maximal 
(resp. minimal) element of $\factor{\sigma}{i}\factor{\sigma}{i-1}$ $\ldots$ $\factor{\sigma}{1}$ is 
$\first{\sigma}{i-1}$.
\end{lemma}

@@

we now define the set
$\SET{\pmotif}{\ptext}{i}{j}$ as follows:
given a subsequence $s$ of $\ptext[j:]$,
$s \in \SET{\pmotif}{\ptext}{i}{j}$ if and only if
$s$ is an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$
and the element $\first{\pmotif}{i}$ occurs in $\ptext[j]$.

\begin{lemma}
Let $\pmotif$ be a permutation , 
$\factor{\pmotif}{i}$ be an ascent (resp. descent) factor,
$s$ be a subsequence such that $s \in \SET{\pmotif}{\ptext}{i}{j}$ and $s[\first{\pmotif}{i-1}]$ is minimal (resp. maximal). 
For all subsequences 
$s' \in \SET{\pmotif}{\ptext}{i}{j}$ and for all subsequences $t$ of $\ptext$, such that $t=t's'$, if $t$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ then the subsequence $t's$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ in $\ptext[j:]$ such that $\first{\pmotif}{i}$ occurs in $\ptext[j]$.
\end{lemma}

@@

\begin{corollary}
Let $\pmotif$ be a permutation , 
$\factor{\pmotif}{i}$ be an ascent (resp. descent) factor
and 
$s$ be a subsequence such that $s \in \SET{\pmotif}{\ptext}{i}{j}$ and $s[\first{\pmotif}{i-1}]$ is minimal (resp. maximal). 
These following statements are equivalent :
\begin{itemize}
	\item There exists an 
	occurrence of $\pmotif$ in $\ptext$ with element $\first{\pmotif}{i}$ occurs in $\ptext[j]$.
	\item There exists a subsequence $t$ of $\ptext[1:j-1]$  such that $ts$ is an occurrence of $\pmotif$ in $\ptext$ with element $\first{\pmotif}{i}$ occurs in $\ptext[j]$. 
\end{itemize}
\end{corollary}

@@

	\text{The maximal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ that is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & 
		\text{ an ascent element } \\	
	\text{ and that minimize the maximal value} & \\
	\text{ and $\first{\sigma}{i}$ occurs in $\ptext[j]$} & \\

@@

	\text{The minimal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ that is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & \text{ a descent element } \\	
	\text{ and that maximize the minimal value} & \\
	\text{ and $\first{\sigma}{i}$ occurs in $\ptext[j]$} & \\
\end{cases}
$$

@@

\noindent\textbf{BASE :}
\begin{align*}
\LM{\pmotif}{\ptext}{1}{j} 
&=
\begin{cases}
		\min_{j<j'}\{\infty\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LIS(j,j',\ptext[j'])  $}\}&\text{ an ascent element } \\
		&\\
		\max_{j<j'}\{0\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LDS(j,j',\ptext[j']) $}\}&\text{ a descent element }\\
\end{cases}
\\
\intertext{
In the base case,
one is looking for an occurrence of the first factor.
If the factor has only ascent (resp. descent) elements, then
one has to find an increasing (resp. decreasing) subsequence
in the text of the same size or that is longer that
the size of the first factor. To find the occurrence that is
the less restrictive (for the future matching)
one must find the occurrence with the right most element
of the sequence minimal (resp. maximal).}
\\
\intertext{\textbf{STEP :}}
\LM{\pmotif}{\ptext}{i}{j} 
&=
\begin{cases}
	\min \{\infty\} \cup  \AF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is an ascent element}\\
	\max \{0\} \cup  \DF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is  a descent element}\\
\end{cases}
\end{align*}
where $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ are the sets 
of elements such that if  
$\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$ or if $\ptext[k] \in \DF{\pmotif}{\ptext}{i}{j}$ 
then there exists an occurrence $t$ of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$ and the element 
$\first{\pmotif}{i-1}$ occurs in element $\ptext[k]$. 
$t$ exists if and only if $t=t's'$ with $t'$ is an occurrence of  $\factor{\pmotif}{i}$ in $\ptext[j:k-1]$ with element
$\first{\pmotif}{i-1}$ occurs in element $\ptext[k]$
and $s \in \SET{\pmotif}{\ptext}{i-1}{k}$.
If $\factor{\pmotif}{i}$ is an increasing (resp. decreasing)
sequence $t'$ must be an increasing (resp. decreasing) 
subsequence
in $\ptext[j:k-1]$ of the same size or that is longer that
the size of $\factor{\pmotif}{i}$ (denoted $|\factor{\pmotif}{i}|$).
Moreover one must assure that the elements of subsequence 
$t'$ are smaller (resp. bigger) than the element of $s'$ 
in particular the minimal (resp. maximal) element.
Thus we can compute $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ as follows:
\begin{align*}
\AF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq 0$ and} 
\\
&\qquad 
\text{$|\factor{\pmotif}{i}| \leq LIS_{\ptext}(j,j'-1,\LM{\pmotif}{\ptext}{i-1}{j'})\}$} 
\\
\DF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq \infty$ and}
\\
&\qquad
\text{$|\factor{\pmotif}{i}| \leq LDS_{\ptext}(j,j'-1,\LM{\pmotif}{\ptext}{i-1}{j'})\}$}
\end{align*}
where $LIS_{\ptext}(i,j,k)$ (resp. $LDS_{\ptext}(i,j,k)$) is the longest increasing 
(resp. decreasing) sequence in $\ptext$ starting at $i$ and ending at $j$,
with every element of this sequence
smaller (resp. bigger) than $k$.
$LIS_{\ptext}$ and $LDS_{\ptext}$ can be computed in 
$O(|\ptext|^2\log(\log(|\ptext|)))$ time (see \cite{Bespamyatnikh00enumeratinglongest}).

There exists a occurrence of $\pmotif$ in $\ptext$ if and only if
there exists a $LM(n,i)\neq 0$ and $LM(n,i)\neq \infty$ for $1 \leq i \leq |\ptext|$,
with $n$ the number of factors in $\pmotif$.
Moreover the basic case can be computed in $O(n^2)$ time
and the induction on $O(kn^2)$ time and space.
\qed
\end{proof}

@@

This section is devoted to the pattern matching problem with $(213,231)$-avoiding bivincular motif.
Recall that a bivincular motif generalise a permutation motif by
be able to force elements to be consecutive in value or in position.
Hence,  bivincular motif is stronger in constraint that motif by permutation, intuitively we can not use the previous algorithm.
As in a $(213,231)$-avoiding permutation, we can describe structural property of a $(213,231)$-avoiding bivincular motif.

\begin{lemma}
\label{lemma:ascentDescentAscent}
Given $\pmotif$ a $(213,231)$-avoiding bivincular,
If $\overline{\sigma[i]\sigma[j]}$ (this implies that $\sigma[j]=\sigma[i]+1$) and 
if $\sigma[i]$ is an ascent (resp. decent) element and $\sigma[i]+1$ is an ascent 
(resp. decent) element then 
$i<j$ (resp. $j>i$) and for every $k$, $i<k<j$ (resp. $j>k>i$), $\pi[k]$ is a descent 
(resp. ascent) element.    
\end{lemma}

A way to understand this lemma is if two ascent (resp. decsent) elements
need to occur in consecutive elements in value then every element in $\pmotif$ which are between those two elements (if any) are descent (resp. ascent) elements.

\begin{proof}
Suppose that there exists $k$, $i<k<j$,
such that $\pmotif[k]$ is ascent. Ascent elements are increasing  so $\sigma[i]<\sigma[k]<\sigma[j]$ which is in contradiction with $\sigma[j]=\sigma[i]+1$.
We use a similar argument if $\sigma[i]$ is a descent element
\qed
\end{proof}

\begin{proposition}
\label{Proposition:bivincular pattern}
Let $\sigma$ be a $(213,231)$-avoiding bivincular of length $k$
and $\pi \in S_n$.
One can decide in $O(kn^5)$ time
and $O(kn^5)$ space if $\sigma$ occurs in $\pi$.
\end{proposition}

\begin{proof}
We consider the following problem :
Given $\pmotif$ a $(213,231)$-avoiding bivincular, and a text $\ptext \in \Av{n_\ptext}{231,213}$,
$$
\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}=
\begin{cases}
	true 	& \text{If $\pbmotif[i:]$ occurs in $\ptext[j:]$}\\
			& \text{with every element is in $[\lb,\ub]$}\\
			& \text{and element $\pmotif[i]$ occurs in $\ptext[j]$ }\\
	false 	& otherwise\\
\end{cases}
$$

$\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}$ can be solved by means by the following relations:

@@

The base case finds an occurrence for the rightmost element of the motif. If the last element does not have any restriction on positions and on values, then $\PM{\pbmotif}{\ptext}{\lb}{\ub}{n_\pmotif}{j}$ is true if and only if $\pmotif[n_\pmotif]$ occurs in $\ptext[j]$. This is true if $\ptext[j] \in [\lb,\ub]$.

@@

We need to consider 3 cases for the problem $\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}$ :
\begin{itemize}
	\item If $\ptext[j] \notin [\lb,\ub]$ then :
	$$
	\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j} = false
	$$

	which is immediate from the definition. If $\ptext[j] \notin [\lb,\ub]$ then it can not be part of an occurrence of  $\pmotif[i:]$ in $\ptext[i:]$ with every element is in $[\lb,\ub]$.

@@

	\end{cases}
	$$	
	Remark that $\pmotif[i]$ occurs in $\ptext[j]$ because $\ptext[j] \in [\lb,\ub]$.
	Thus if $\pmotif[i+1:]$ occurs in $\ptext[j+1:]$ with every element in $[\pmotif[i]+1,\ub]$ then 
 	$\pmotif[i:]$ occurs in $\ptext[j:]$  	
 	which correspond to know if there exists $k$, $j<k$ such that
	$\PM{\pbmotif}{\ptext}{\ptext[j]+1}{\ub}{i+1}{k}$ is true.

@@

	$$
	\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}=
	\begin{cases}
			\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}
				& \text{if $\pmotif[i]$ is not underlined } \\
				& \text{ and $\pmotif[i]$ is not overlined} \\
			\bigcup_{j<k} \PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{k}
				& \text{if $\pmotif[i]$ is not underlined } \\
				& \text{ and $\overline{\pmotif[i](\pmotif[i]+1)}$ or ${\pmotif[i]}^\urcorner$}\\
				& \text{ and $\ptext[j]=\ub$} \\
			\PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{j+1}
				& \text{if $\underline{\pmotif[i]\pmotif[i+1]}$ } \\
				& \text{ and $\overline{\pmotif[i](\pmotif[i]+1)}$ or ${\pmotif[i]}^\urcorner$}\\
				& \text{ and $\ptext[j]=\ub$} \\
			\PM{\pbmotif}{\ptext}{\lb}{\ptext[j]-1}{i+1}{j+1}
				& \text{if $\underline{\pmotif[i]\pmotif[i+1]}$ } \\
				& \text{ and $\pmotif[i]$ is not overlined} \\
			false & \text{otherwise}
	\end{cases}
	$$	
	The same remark as the last case holds.

@@

\begin{proposition}
Given two permutations $\pi_1$ and $\pi_2$,
one can compute 
the longest common subsequence
avoiding (213,231) in $O(|\pi_1|^3|\pi_2|^3)$ time and space.
\end{proposition}
